The Talking Heads Experiment

### 2015-04-23 chap8/chap8.tex
hyphenation, punctuation, positioning, etc for chapters 1-6
hyphenations, footnotes, ...

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap8/figures/th-website.pdf}}
\caption{\label{fig:thwebsite}The Talking Heads website main interface. Users could create and manage their own agents and inspect what was happening 
at each site. Here 12 agents located at the Brussels site are listed.}
\end{figure}

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.85\textwidth]{chap8/figures/zelebot.pdf}}
\caption{\label{fig:thwebsite-agent}Example of an interaction that took part in Antwerp. This game was a success. The agents recognised the same
set of objects and the meaning of `kazozo' was effective in finding the same topic. 
}
\end{figure}

@@

Once the physical installation at the Antwerp site was completed, work focused on getting the experiment itself 
up and running. It is interesting and instructive to look at some snippets of the email correspondence during the 
first weeks. The despair when trying to cope with the unavoidable problems but also the excitement 
as a language began to emerge shines through. The English humor of Angus McIntyre was a welcome antidote to 
all the stress. The correspondence reprinted here is a small selection but also necessary fragmentary because a 
lot of communication took place face-to-face or through the telephone, as email was not always possible. 
\enlargethispage{1\baselineskip}
We pick up the email correspondence on
June 23, 1999 when the first experiments are taking place to try and make language games at the three sites possible 
and allow the scheduling and activation of agents.

\subsection*{23 June 1999}

\begin{mail}
--============\_-1278188520==\_d============\\
From: Angus McIntyre <angus@csl.sony.fr>\\
Subject: Re: Brussels is back\\

@@

I've announced the site to a handful of links pages (Peter Norvig's 
AI links collection, Chris Bogart's Constructed Languages page, the 
NASA real robots page - links to all these sites are under 
'background' on our site). I haven't yet done a massive submit of 
the site to the big search engines or announcement sites; I suggest 
that we let traffic build slowly so we can find out what we can handle.
	A
\end{mail}

@@

The (faster) machines that were planned are now available, so that a switch has to be made while the 
experiment is already running. 
\newline
{
\begin{mail}
--============\_-1278188520==\_d============\\
From: Angus McIntyre <angus@csl.sony.fr>\\
Subject: 'headnet' now running on new PII 400 server.\\

@@

I'm off home now. If you need me to come back and kick 
the server, call me on:
	+33-1-42-78-73-88
Bye,
	A
\end{mail}
}
\clearpage
\subsection*{25 June 1999}

>>>>

\begin{figure}[htb]
  \centerline{\includegraphics[width=.85\textwidth]{chap8/figures/th-website.pdf}}
\caption{\label{fig:thwebsite}The Talking Heads website main interface. Users could create and manage their own agents and inspect what was happening 
at each site. Here 12 agents located at the Brussels site are listed.}
\end{figure}

@@

\begin{figure}[htb]
  \centerline{\includegraphics[width=.85\textwidth]{chap8/figures/zelebot.pdf}}
\caption{\label{fig:thwebsite-agent}Example of an interaction that took part in Antwerp. This game was a success. The agents recognised the same
set of objects and the meaning of `kazozo' was effective in finding the same topic. 
}
\end{figure}

@@

Once the physical installation at the Antwerp site was completed, work focused on getting the experiment itself 
up and running. It is interesting and instructive to look at some snippets of the email correspondence during the 
first weeks. The despair when trying to cope with the unavoidable problems but also the excitement 
as a language began to emerge shines through. The English humor of Angus McIntyre was a welcome antidote to 
all the stress. The correspondence reprinted here is a small selection but also necessary fragmentary because a 
lot of communication took place face-to-face or through the telephone, as email was not always possible. 
\clearpage
We pick up the email correspondence on
June 23, 1999 when the first experiments are taking place to try and make language games at the three sites possible 
and allow the scheduling and activation of agents.

\subsection*{23 June 1999}
\begin{mail}
--============\_-1278188520==\_d============\\
From: Angus McIntyre <angus@csl.sony.fr>\\
Subject: Re: Brussels is back\\

@@

I've announced the site to a handful of links pages (Peter Norvig's 
AI links collection, Chris Bogart's Constructed Languages page, the 
NASA real robots page - links to all these sites are under 
'background' on our site). I haven't yet done a massive submit of 
the site to the big search engines or\clearpage announcement sites; I suggest 
that we let traffic build slowly so we can find out what we can handle.
	A
\end{mail}

@@

The (faster) machines that were planned are now available, so that a switch has to be made while the 
experiment is already running. 
\newline

\begin{mail}
--============\_-1278188520==\_d============\\
From: Angus McIntyre <angus@csl.sony.fr>\\
Subject: 'headnet' now running on new PII 400 server.\\

@@

I'm off home now. If you need me to come back and kick 
the server, call me on:
	+33-1-42-78-73-88
Bye,
	A
\end{mail}

\clearpage
\subsection*{25 June 1999}

### 2015-04-23 chap9/chap9.tex
hyphenations, footnotes, ...

To express conceptualisations, agents need a lexicon relating words to meanings. It must function 
bi-directionally (words-to-meanings and meanings-to-words); it must also store synonyms (more than one word for the same meaning) and ambiguities (more than one meaning for the same word). The agents have not been given a lexicon; they must acquire their own common lexicon as a bi-product of the game.

@@

A website has been created for the experiment\footnote{http://talking-heads.csl.sony.fr/}. Through this site, anybody who wants can create new agents and follow their progress. Owners of agents can teach them words, so that words already used in human natural languages sneak into the emerging vocabulary. The agents are autonomous and do not necessarily stick to the words given to them but try to maximally adapt to the behavior of the group and invent their own words. Through the website it is also possible to monitor the progress of the experiment: the lexicon being created, the success rate, the coherence among the agents, the complexity of the language, etc. There is a Hall of Fame listing the best speakers and hearers. This motivates humans to take care of their agents thus ensuring that they move to the top in the Hall of Fame.

To express a conceptualisation, agents need a lexicon relating words with meanings. The lexicon must be consultable in both directions (from words to meanings and from meanings to words). It must be able to store synonyms (more than one word for the same meaning) and ambiguities The agents have not been given a lexicon a priori. They have to acquire their own lexicon as a side effect of the game. New words get into a lexicon in two ways: When an agent has no word to express a particular distinction, the agent can create a new one by a random combination of syllables. When an agent hears a word that he does not know, he stores the new word with his own guess of what the meaning could be in the scene being perceived.

@@

{\bfshape 2000-02-26 20:41:20 Trash}\\
{\itshape RE: teaching / TRIBES}	\\
Basically think about it like this... On the Earth there are many different languages spoken by people usually in a specific geographical location. Now myself and yeahbaby looked at this experiment (as we are very interested in AI and have done much work in this field) and decided to test the idea of ``Tribes" where all those in that tribe speak the same lingo. We also figured that tribes rarely move out of their ``Birth Location" so kept them at one site. As can be seen now ALL our Agents speak the same language and have even taught it to others outside the Tribe (such as the esteemed Yellow Mk 1 Escort  and possibly even Anne's little Agent). 
These agents we considert to be forreigners!! But stilll they can communicate with our Agents... If you have noticed our words in the lexicon as explained by yeahbaby have  very few deviations from what was intended unlike others such as ``Gumble" which appears to mean Everything!!!!!

@@

{\bfshape 2000-02-27 19:52:20	Trash}\\
{\itshape RE: Collusion Tribes}\\
At the minute me and Yeah are off the Paris server (we are regrouping our thoughts for further action!!!)
But yes I did mean view the pictures when i said definition.\\
By 2 tribes I mean obviously two who have been taught on the same server.
So say that i train my agents on there this week and you train your tribe next week on the same server with different words..... when both tribes are capable of good communication we then take half of each tribe and place them together in the 3rd week on the Paris server and see if they become multi lingual.\\
We then sned these linguists back to their own tribe and see what happens.\\
Unfortunately many people have tried to jump on the band wagon of Paris success and send their Agents there with no idea what they are doing ... this has meant that we can not get our tribe to dominate as we would like.....\\
The problems with organising tribe transfer is immense and it is a shame people don't exp[lain what they are doing on this fprum so that all could help out by say teaching their agents the words being used or even evacuating a site so that two tribes can collide ...\\
It is a real shame that an experiment into AI and Language is hampered because people are unwilling to communicate with each other .... \\
Well I guess that those of us who want a proper experiment should just ry their best in difficult circumstances...\\
About the wrong words being taught !!! If your Tribe is more than 60\% of the server then your tribe generallly succeeds in teaching the ``foreigners'' with little effect on themselves ... This has more to do with Probability than anything else.

@@

{\bfshape 2000-02-27 20:46:29 smartypants}\\
{\itshape RE: Collusion Tribes}\\
Thanks Trash. I was thinking in different terms...\\
... seeing as anyone doing anything is on the Paris server, I was imagining that as one `big' tribe and that any outsiders (stuck on the other servers, or new) are the `other' tribes so thanks for clearing that up for me.\\
I'd be interested in working with you and Yeahbaby for a more meaningful experiment, particularly as loads of `dumber' robots (by that I mean ones not enjoying the Paris success) will soon be unleashed on us.\\
Just to clarify, my plan up until now was to let my robots learn 40-50 words (as suggested in the tips), I started to teach them the most successful words from the lexicon yesterday (most of which were yours - grrrrrr, jealous!) and then I sent them to Paris. (This was before I started putting stuff up on the forum and saw your theory unfortunately). \\
The only one that got in was Virtuoso, and even then he didn't really get to play enough games as I was experimenting to see if the number of games you choose has any effect on getting into that server.\\
Anwyay, let me know if you're interested (I've only created 7 at the moment).\\
By the way, do I remember seeing somewhere that you're both girls?????\\
... `cause I am...\\
If so it is interesting to see different ways people approach the experiment. If I was running the experiment I reckon I could do research from the forum, as well as the actual experiment.\\

@@

To deal with the problem of agent congestion, software issues, and camera misalignment at the London 
and Cambridge sites, maintenance was carried out which lead to a temporary unavailability of the experiment. 
Surprisingly, the reaction of the Hull group was entirely negative as they saw it as a 
way to prevent them from gaining or keeping control of the language used by the agents, even though 
Angus McIntyre clearly and patiently explained why maintenance was necessary. Also, Denna Jones ({\it Phlox}) took 
\enlargethispage{1em}
it personally as she did not realise that the calibration errors were due to vibrations caused by
traffic of the London Underground near and under the Wellcome Gallery.

@@

\newpage
\begin{mail}

@@

{\bfshape 2000-03-08 15:33:00 fish}\\
{\itshape RE: Hull5 Sony 0}\\
big hand to the hull boys.
you may be a bit sensitive to personal insults but atleast you can kick ass when it needs it.....\\
already seen what you're doing in brussels.... and we love it!\\

@@

{\bfshape 2000-03-08 17:31:27 phlox}\\
{\itshape RE: Phloxy Lady}\\
too right. do it.\\

@@

{\bfshape 2000-03-09 16:13:23 phlox}\\
{\itshape we're still waiting. . .}\\
come on sony own up. if youre gonna exercise stalinistic control at least do it up front; not under the guise of testing software;. the only agents on site at the mo are owned by luc, angus, adam and joris. if it's in the public domain - and it is -  then let the public play!
%
\end{mail}
\clearpage

@@

The Look into the Box installation was shown again at several locations. One was in July 2003  
in the context of a yearly music festival in Spoleto. In 2003, the theme of semiotic dynamics was 
chosen and various presentations and 
discussions were held curated by the Italian semiotician Paolo Fabbri. A new installation of 
Look into the Box was realised and operated (see \figref{fig:spoleto}). A system for playing language games 
with human users, created by Tony Belpaeme of the VUB AI Lab, was also demonstrated. This system was intended
to collect data about colour category prototypes and names from human speakers in order to gather more data
about colour language in discrimination games.

>>>>

To express conceptualisations, agents need a lexicon relating words to meanings. It must function 
bi-directionally (words-to-meanings and meanings-to-\linebreak[4] words); it must also store synonyms (more than one word for the same meaning) and ambiguities (more than one meaning for the same word). The agents have not been given a lexicon; they must acquire their own common lexicon as a bi-product of the game.

@@

A website has been created for the experiment\footnote{http://talking-heads.csl.sony.fr/}. Through this site, anybody who wants can create new agents and follow their progress. Owners of agents can teach them words, so that words already used in human natural languages sneak into the emerging vocabulary. The agents are autonomous and do not necessarily stick to the words given to them but try to maximally adapt to the behavior of the group and invent their own words. Through the website it is also possible to monitor the progress of the experiment: the lexicon being created, the success rate, the coherence among the agents, the complexity of the language, etc. There is a Hall of Fame listing the best speakers and hearers. This motivates humans to take care of their agents thus ensuring that they move to the top in the Hall of Fame.
\clearpage
To express a conceptualisation, agents need a lexicon relating words with meanings. The lexicon must be consultable in both directions (from words to meanings and from meanings to words). It must be able to store synonyms (more than one word for the same meaning) and ambiguities The agents have not been given a lexicon a priori. They have to acquire their own lexicon as a side effect of the game. New words get into a lexicon in two ways: When an agent has no word to express a particular distinction, the agent can create a new one by a random combination of syllables. When an agent hears a word that he does not know, he stores the new word with his own guess of what the meaning could be in the scene being perceived.

@@

\rule{0.8\textwidth}{.4pt}
\clearpage
{\bfshape 2000-02-26 20:41:20 Trash}\\
{\itshape RE: teaching / TRIBES}	\\
Basically think about it like this... On the Earth there are many different languages spoken by people usually in a specific geographical location. Now myself and yeahbaby looked at this experiment (as we are very interested in AI and have done much work in this field) and decided to test the idea of ``Tribes" where all those in that tribe speak the same lingo. We also figured that tribes rarely move out of their ``Birth Location" so kept them at one site. As can be seen now ALL our Agents speak the same language and have even taught it to others outside the Tribe (such as the esteemed Yellow Mk 1 Escort  and possibly even Anne's little Agent). 
These agents we considert to be forreigners!! But stilll they can communicate with our Agents... If you have noticed our words in the lexicon as explained by yeahbaby have  very few deviations from what was intended unlike others such as ``Gumble" which appears to mean Everything!!!!!

@@

{\bfshape 2000-02-27 19:52:20	Trash}\\
{\itshape RE: Collusion Tribes}\\
At the minute me and Yeah are off the Paris server (we are regrouping our thoughts for further action!!!)
But yes I did mean view the pictures when i said definition.\\
By 2 tribes I mean obviously two who have been taught on the same server.
So say that i train my agents on there this week and you train your tribe next week on the same server with different words..... when both tribes are capable of good communication we then take half of each tribe and place them together in the 3rd week on the Paris server and see if they become multi lingual.\\
We then sned these linguists back to their own tribe and see what happens.\\ \clearpage
Unfortunately many people have tried to jump on the band wagon of Paris success and send their Agents there with no idea what they are doing ... this has meant that we can not get our tribe to dominate as we would like.....\\
The problems with organising tribe transfer is immense and it is a shame people don't exp[lain what they are doing on this fprum so that all could help out by say teaching their agents the words being used or even evacuating a site so that two tribes can collide ...\\
It is a real shame that an experiment into AI and Language is hampered because people are unwilling to communicate with each other .... \\
Well I guess that those of us who want a proper experiment should just ry their best in difficult circumstances...\\
About the wrong words being taught !!! If your Tribe is more than 60\% of the server then your tribe generallly succeeds in teaching the ``foreigners'' with little effect on themselves ... This has more to do with Probability than anything else.

@@

{\bfshape 2000-02-27 20:46:29 smartypants}\\
{\itshape RE: Collusion Tribes}\\
Thanks Trash. I was thinking in different terms...\\
... seeing as anyone doing anything is on the Paris server, I was imagining that as one `big' tribe and that any outsiders (stuck on the other servers, or new) are the `other' tribes so thanks for clearing that up for me.\\
I'd be interested in working with you and Yeahbaby for a more meaningful experiment, particularly as loads of `dumber' robots (by that I mean ones not enjoying the Paris success) will soon be unleashed on us.\\
Just to clarify, my plan up until now was to let my robots learn 40-50 words (as suggested in the tips), I started to teach them the most successful words from the lexicon yesterday (most of which were yours - grrrrrr, jealous!) and then I sent them to Paris. (This was before I started putting stuff up on the forum and saw your theory unfortunately). \\
The only one that got in was Virtuoso, and even then he didn't really get to play enough games as I was experimenting to see if the number of games you choose has any effect on getting into that server.\\
Anwyay, let me know if you're interested (I've only created 7 at the moment).\\
By the way, do I remember seeing somewhere that you're both girls?????\\
... `cause I am...\clearpage
If so it is interesting to see different ways people approach the experiment. If I was running the experiment I reckon I could do research from the forum, as well as the actual experiment.\\

@@

To deal with the problem of agent congestion, software issues, and camera misalignment at the London 
and Cambridge sites, maintenance was carried out which lead to a temporary unavailability of the experiment. 
Surprisingly, the reaction of the Hull group was entirely negative as they saw it as a 
way to prevent them from gaining or keeping control of the language used by the agents, even though 
Angus McIntyre clearly and patiently explained why maintenance was necessary. Also, Denna Jones ({\it Phlox}) took 
it personally as she did not realise that the calibration errors were due to vibrations caused by
traffic of the London Underground near and under the Wellcome Gallery.

@@

@@

{\bfshape 2000-03-08 15:33:00 fish}\\
{\itshape RE: Hull5 Sony 0}\\
big hand to the hull boys.
you may be a bit sensitive to personal insults but atleast you can kick ass when it needs it.....\\ \enlargethispage{1\baselineskip}
already seen what you're doing in brussels.... and we love it!\\

@@

\rule{0.8\textwidth}{.4pt}
\clearpage
{\bfshape 2000-03-08 17:31:27 phlox}\\
{\itshape RE: Phloxy Lady}\\
too right. do it.\\

@@

{\bfshape 2000-03-09 16:13:23 phlox}\\
{\itshape we're still waiting. . .}\\
come on sony own up. if youre gonna exercise stalinistic control at least do it up front; not under the guise of testing software;. the only agents on site at the mo are owned by luc, angus, adam and joris. if it's in the public domain - and it is -  then let the public play!
%
\end{mail}

@@

The Look into the Box installation was shown again at several locations. One was in July 2003  
in the context of a yearly music festival in Spoleto. In 2003, the theme of semiotic dynamics was 
chosen and various presentations and 
discussions were held curated by the Italian semiotician Paolo Fabbri. A new installation of 
Look into the Box was realised and operated (see \figref{fig:spoleto}). A system for playing language games 
with human users, created by Tony Belpaeme of the VUB AI Lab, was also demonstrated. This system was intended
to collect data about colour category prototypes and names from human speakers in order to gather more data \enlargethispage{1\baselineskip}
about colour language in discrimination games.

### 2015-04-23 extra/papers-language-steels.tex
fix capitalization
hyphenations, footnotes, ...

29. \citet{Kaplan:00a} Comment les robots construisent leur monde : Experiences sur la convergence des cat\'egories sensorielles. In: Dessalles, J-L. (ed.), Journ\'e ARC Evolution et Cognition, ENST Paris. pp. 13-18.

@@

51. \citet{Steels:03c} Language-reentrance and the Inner Voice'. Journal of Consciousness Studies, 10(4-5):173-185.

@@

65. \citet{Steels:05d} Linking in Fluid Construction Grammar. Proceedings of BNAIC. Transactions of the Belgian Royal Society for 
Science and Arts. pp. 11--18. October 2005.

@@

75. \citet{Steels:07b} Semiotic Dynamics solves the Symbol Grounding Problem. Nature Precedings 
http://precedings.nature.com/documents/1234/version/1

@@

81. \citet{Steels:07h} Is Symbolic Inheritance Similar to Genetic Inheritance?. Behavioral and Brain Sciences, 30(4):376-377.

@@

112. \citet{Steels:11c} Modeling the cultural evolution of language. Physics of Life Reviews. 8(4) 330-356. http://www.sciencedirect.com/science/article/pii/S157106451100145X

@@

135. \citet{Beuls:13a} Agent-Based Models of Strategies for the Emergence and Evolution of Grammatical Agreement. PLOS ONE, 8(3), e58960. http://dx.plos.org/10.1371/journal.pone.0058960

>>>>

\enlargethispage{1\baselineskip}
29. \citet{Kaplan:00a} Comment les robots construisent leur monde : Experiences sur la convergence des cat\'egories sensorielles. In: Dessalles, J-L. (ed.), Journ\'e ARC Evolution et Cognition, ENST Paris. pp. 13-18.

@@

51. \citet{Steels:03c} Language-reentrance and the Inner Voice. Journal of Consciousness Studies, 10(4-5):173-185.

@@

\enlargethispage{1\baselineskip}
65. \citet{Steels:05d} Linking in Fluid Construction Grammar. Proceedings of BNAIC. Transactions of the Belgian Royal Society for 
Science and Arts. pp. 11--18. October 2005.

@@

75. \citet{Steels:07b} Semiotic Dynamics solves the Symbol Grounding Problem. Nature Precedings 
\url{http://precedings.nature.com/documents/1234/version/1}

@@

\clearpage
81. \citet{Steels:07h} Is Symbolic Inheritance Similar to Genetic Inheritance?. Behavioral and Brain Sciences, 30(4):376-377.

@@

112. \citet{Steels:11c} Modeling the cultural evolution of language. Physics of Life Reviews. 8(4) 330-356. \url{http://www.sciencedirect.com/science/article/pii/S157106451100145X}

@@

135. \citet{Beuls:13a} Agent-Based Models of Strategies for the Emergence and Evolution of Grammatical Agreement. PLOS ONE, 8(3), e58960. \url{http://dx.plos.org/10.1371/journal.pone.0058960}

### 2015-04-23 preface/preface.tex
hyphenations, footnotes, ...

I also thank the curators of the Laboratorium Exhibition in Antwerp (Hans-Ulrich Obrist and Barbara Vanderlinden) 
and the members of its curating board (Bruno Latour and Carsten H\"oller), the curators of the N01SE exhibition 
in London (Adam Lowe and Simon Schaffer), artist Olafur Eliasson\todo{Spelling:\'Olafur El\'iasson ??}, with whom I collaborated on the Look into the Box 
piece at the Mus\'e d'art Moderne in Paris, artist Anne-Mie Van Kerckhoven for collaborations for the Chromosophy 
laboratory in Aachen and the many organisers and helping hands who made the other installations possible. 
Sylvia Spruck Wrigley has helped to improve the 1999 pre-edition (now \partref{part:1} of the book) 
under considerable time pressure and Marleen Wynants was crucial in the last phases of this publication with comments,
professional advice, photographs, and support.

>>>>

I also thank the curators of the Laboratorium Exhibition in Antwerp (Hans-Ulrich Obrist and Barbara Vanderlinden) 
and the members of its curating board (Bruno Latour and Carsten H\"oller), the curators of the N01SE exhibition 
in London (Adam Lowe and Simon Schaffer), artist Olafur Eliasson, with whom I collaborated on the Look into the Box 
piece at the Mus\'e d'art Moderne in Paris, artist Anne-Mie Van Kerckhoven for collaborations for the Chromosophy 
laboratory in Aachen and the many organisers and helping hands who made the other installations possible. 
Sylvia Spruck Wrigley has helped to improve the 1999 pre-edition (now \partref{part:1} of the book) 
under considerable time pressure and Marleen Wynants was crucial in the last phases of this publication with comments,
professional advice, photographs, and support.
